Arrows for Parallel Computation

### 2018-01-06 paper/content/abstract.tex
phil first email remarks
minor things

\begin{abstract}
Arrows are a general interface for computation and an alternative to Monads for API design. In contrast to Monad based parallelism, we explore the use of Arrows for specifying generalised parallelism. Specifically, we define an Arrow-based language and implement it using multiple parallel Haskells.

@@

To show that PArrows have similar expressive power as existing parallel languages, we implement several skeletons and 4 benchmarks. 
Benchmarks show that our framework does not induce any notable performance overhead. We conclude that Arrows have considerable potential for composing parallel programs, and for producing programs that can execute on multiple parallel language implementations.

>>>>

\begin{abstract}
Arrows are a general interface for computation and an alternative to Monads for API design. In contrast to Monad-based parallelism, we explore the use of Arrows for specifying generalised parallelism. Specifically, we define an Arrow-based language and implement it using multiple parallel Haskells.

@@

To show that PArrows have similar expressive power as existing parallel languages, we implement several skeletons and four benchmarks. 
Benchmarks show that our framework does not induce any notable performance overhead. We conclude that Arrows have considerable potential for composing parallel programs, and for producing programs that can execute on multiple parallel language implementations.

### 2018-01-06 paper/content/motivation.tex
minor things

\paragraph{Contributions}
%
%\olcomment{HIT HERE REALLY STRONG}
%
%\subsection{Impact of parallel Arrows}
%\olcomment{move this to Contributions in the front or something}
We propose an Arrow-based encoding for parallelism based on a new Arrow combinator |parEvalN :: [arr a b] -> arr [a] [b]|. A parallel Arrow is still an Arrow, hence the resulting parallel Arrow can still be used in the same way as a potential sequential version. In this paper we evaluate the expressive power of such a formalism in the context of parallel programming.

@@

This has many practical advantages. For example, during development we can run the program in a simple GHC-compiled variant using a GpH backend and afterwards deploy it on a cluster by converting it into an Eden program, by just replacing the |ArrowParallel| instance and compiling with Eden's GHC variant. (Section~\ref{sec:parallel-arrows})
\item We extend the PArrows formalism with |Future|s to enable direct communication of data between nodes in a distributed memory setting similar to Eden's Remote Data (|RD|) \citet{Dieterle2010}. Direct communication is useful in a distributed memory setting because it allows for inter-node communication without blocking the master-node. (Section~\ref{sec:futures})
\item We demonstrate the expressiveness of PArrows by using them to define common algorithmic skeletons ((Section~\ref{sec:skeletons}), and by using these skeletons to implement four benchmarks (Section~\ref{sec:benchmarks}).
\item We practically demonstrate that Arrow parallelism has a low performance overhead compared with existing approaches, \eg with mean of mean overhead of less than $3.5\%, 0.8\%$ for all benchmarks with GpH and Eden respectively. Compared to |Par| Monad we virtually noticed no overhead with a mean of mean overheads of less than $0\%$ in all benchmarks. (Section~\ref{sec:benchmarks}).
\end{itemize}

PArrows are open source and available from \url{https://github.com/s4ke/Parrows}.

>>>>

\paragraph{Contributions.}
%
%\olcomment{HIT HERE REALLY STRONG}
%
%\subsection{Impact of parallel Arrows}
%\olcomment{move this to Contributions in the front or something}
We propose an Arrow-based encoding for parallelism based on a new Arrow combinator |parEvalN :: [arr a b] -> arr [a] [b]|. A parallel Arrow is still an Arrow, hence the resulting parallel Arrow can still be used in the same way as a potential sequential version. In this paper we evaluate the expressive power of such a formalism in the context of parallel programming.

@@

This has many practical advantages. For example, during development we can run the program in a simple GHC-compiled variant using a GpH backend and afterwards deploy it on a cluster by converting it into an Eden program, by just replacing the |ArrowParallel| instance and compiling with Eden's GHC variant. (Section~\ref{sec:parallel-arrows})
\item We extend the PArrows formalism with |Future|s to enable direct communication of data between nodes in a distributed memory setting similar to Eden's Remote Data \citep[|RD|,][]{Dieterle2010}. Direct communication is useful in a distributed memory setting because it allows for inter-node communication without blocking the master-node. (Section~\ref{sec:futures})
\item We demonstrate the expressiveness of PArrows by using them to define common algorithmic skeletons ((Section~\ref{sec:skeletons}), and by using these skeletons to implement four benchmarks (Section~\ref{sec:benchmarks}).
\item We practically demonstrate that Arrow parallelism has a low performance overhead compared with existing approaches, \eg with mean of mean overhead of less than $3.5\%$ and less than $0.8\%$ for all benchmarks with GpH and Eden, respectively. Compared to |Par| Monad we noticed no overhead: the mean of mean overheads was in our favour in all benchmarks (Section~\ref{sec:benchmarks}).
\end{itemize}

PArrows are open source and are available from \url{https://github.com/s4ke/Parrows}.

### 2018-01-06 paper/content/relwork.tex
minor things

\paragraph{Arrows in other languages.}
Although this work is centred on Haskell implementation of Arrows, it is applicable to any functional programming language where parallel evaluation and Arrows can be defined. Basic definitions of PArrows are possible in the Frege language\footnote{GitHub project page at \url{https://github.com/Frege/frege}} (which is basically Haskell on the JVM). However, they are beyond the scope of this work, as are similar experiments with the Eta language\footnote{Eta project page at \url{eta-lang.org}}, 
a new approach to Haskell on the JVM.

>>>>

\paragraph{Arrows in other languages.}
Although this work is centred on Haskell implementation of Arrows, it is applicable to any functional programming language where parallel evaluation and Arrows can be defined. Basic definitions of PArrows are possible in the Frege language\footnote{GitHub project page at \url{https://github.com/Frege/frege}} (which is basically Haskell on the JVM). However, they are beyond the scope of this work, as are similar experiments with the Eta language\footnote{Eta project page at \url{http://eta-lang.org}}, 
a new approach to Haskell on the JVM.
